---
title: "time series intro"
author: "Shelby Walsh"
date: "1/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(lubridate)
library(tsibble)
library(feasts)
library(slider)
```

## Read in data
```{r}
toolik <- read_csv(here("data", "toolikweather.csv"))
```

### Convert to time series data
```{r}
toolik_ts <- toolik %>% 
  mutate(date = lubridate::mdy(date)) %>% 
  as_tsibble(key = NULL, index = date)


```

```{r}
ggplot(data = toolik_ts, aes(x = date, y = mean_airtemp)) +
  geom_line()
```

## Use Index_by function to aggregate time series info by intercals we specify

```{r}

toolik_month <- toolik_ts %>% 
  index_by(yr_mo = ~yearmonth(.)) %>% 
  summarize(monthly_mean_temp = mean(mean_airtemp, na.rm = TRUE))

ggplot(data = toolik_month, aes(x = yr_mo, y = monthly_mean_temp)) +
  geom_line() +
  facet_wrap(~month(yr_mo, label = TRUE))
```

### Aggregate by week
```{r}
toolik_weekly <- toolik_ts %>% 
  index_by(weekly = ~yearweek(.)) %>% 
  summarize(weekly_airtemp = mean(mean_airtemp, na.rm = TRUE),
            min_airtemp = min(mean_airtemp))

head(toolik_weekly)
```

## Filtering time series

use filter_indes to specify ranges of ts to keep or explore
```{r}
toolik_ts %>% 
  filter_index("2000-06" ~ "2001-10")
```

```{r}
toolik_ts %>% 
  filter_index("2006-04-10" ~ "2006-05-15")
```
## Seasonplots
```{r}
toolik_ts %>% 
  filter(year(date) > 2014) %>% 
  gg_season(y = mean_airtemp)

toolik_month %>% 
  gg_season(y = monthly_mean_temp) +
  theme_dark()
```

## seasonal subseries plot

```{r}
toolik_month %>% 
  gg_subseries(monthly_mean_temp)
```

# window: plotting averagge of windows

## Find moving average, use slider package

```{r}
set.seed(2021)
test <- rnorm(n =100, mean = 40, sd = 10)

slide(test, ~.x, .before = 2, .afer = 2)
```


```{r}
roll_toolik_15 <- toolik_ts %>% 
  mutate(ma_15d = as.numeric(slide(toolik_ts$mean_airtemp,
                                   mean,
                                   .before = 7,
                                   .after = 7)))

ggplot() +
  geom_line(data = toolik_ts, aes(x = date, y = mean_airtemp),
            size = 0.2,
            color = "gray") +
  geom_line(data = roll_toolik_15, aes(x = date, y = ma_15d)) +
  theme_minimal()
```

### Create the ACF

```{r}
toolik_ts %>% 
  ACF(mean_airtemp) %>% 
  autoplot()

toolik_month %>% 
  ACF(monthly_mean_temp) %>% 
  autoplot()
```

## Smoothing decomposition by STL

```{r}
toolik_dec <- toolik_month %>% 
  model(STL(monthly_mean_temp ~ season(window = Inf)))

components(toolik_dec) %>% autoplot()
```

